A hierarchical approach for modelling X-ray beamlines. Application to a coherent beamline.

### 2019-06-17 paper-hierarchical.tex
update
arxiv almost ready

\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{color,xcolor}
\usepackage{siunitx}
\usepackage[normalem]{ulem}

@@

\begin{abstract}
We consider different approaches to simulating a modern X-ray beamline. Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performance. Parameters such as flux, dimensions and intensity distribution of the focused beam and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray-tracing and wave optics techniques. A latest-generation X-ray nanofocusing beamline for coherent applications (ID16A at the ESRF) has been chosen to study in detail the issues related to highly demagnifying synchrotron sources and exploiting the beam coherence. The performance of the beamline is studied for two storage rings: the old ESRF-1 (emittance 4000~pm) and the new ESRF-EBS (emittance 150~pm). In addition to traditional results in terms of flux and beam sizes, an innovative study on the partial coherence properties based on the propagation of coherent modes is presented. The different algorithms and methodologies are implemented in the software suite OASYS \cite{codeOASYS}. Those are discussed with emphasis placed upon the their benefits and limitations.
\end{abstract}

@@

Accurate calculation and quantitative evaluation of the parameters related to X-ray optics are of great importance for designing, building and exploiting new beamlines. Every modern beamline follows a procedure of conceptual design plus technical design where detailed simulations are essential. These calculations allowing testing of the design parameters in a virtual computer environment where advantages and limitations can be studied quantitatively. Selection of the optimal optical layout requires the detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optical configuration. Deformations of the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformations usually done using finite element analysis. Traditionally the optics calculations for synchrotron beamlines are done using ray-tracing techniques. They model the X-ray beam as a collection of incoherent rays. This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc. However, the high coherence of the new sources can introduce the necessity to complement the ray-tracing methods with models including effects of diffraction and scattering from a coherent beam. A fully coherent hard X-ray beam is still a dream for the new generation storage rings. Despite a considerable increase in the coherent fraction of the new beams, the coherent fraction at typically hard X-ray photon energies is still a few percents for the new sources. Ray-tracing analysis does not become obsolete but has to be complemented with an analysis based on wave optics. Because of the partial coherence, the wave optics analysis becomes nontrivial and can be done with different levels of approximation, as discussed in this work. In addition, wave optics simulations are usually computationally expensive.

To illustrate the applicability of a variety of methods and solutions for beamline optical simulations we have chosen the ESRF beamline ID16A~\cite{ID16A}. This beamline provides a high-brilliance beam focused to a few nanometers. It displays characteristics increasingly aimed for in modern beamlines: extremely high demagnification (sub-micrometer focused beam size) and high coherence. It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state~\ingreen{\cite{XRFID16A}}.

@@

 In this paper, some quantitative results on how the storage ring upgrade will affect a particular beamline are presented and it is structured as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and explain the particular choice of the parameters used in the calculations. Section~\ref{Simulations} is the central part of the paper, and describes the different methods used to analyze the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline. In Section \ref{level0} we calculate with simple methods approximate values of focal size, flux and coherent fraction. Accuracy can be improved using geometrical methods in Section \ref{level1} that allow a more complete simulation of the beamline optics within certain limits which are related, in this example, to high demagnification and high coherence of the studied beamline. The wave optics section~\ref{level2} includes different models of increasing complexity as the required accuracy of the results becomes more demanding. In the Discussion (Section \ref{discussion}) the results of the different simulation methods are commented. The hierarchical approach allows more accurate and better results but the price to pay is to deal with more complex and computationally expensive tools. Finally, Section \ref{summary} summarizes the results.

@@

The undulator natural photon source size and divergence are given by \cite{elleaume}:
\begin{align}
    \label{eq:photon small sigmas}
    \sigma_u=\frac{2.74}{4\pi}\sqrt{\lambda L}\approx \frac{1}{2  \pi}\sqrt{2 \lambda L}  && \sigma_u' = 0.69\sqrt{\frac{\lambda}{L}}\approx \sqrt{\frac{\lambda}{2 L}},
\end{align}
where $\lambda$ is the emitted wavelength and $L=N\Lambda_u$ is the undulator length, with $\Lambda_u$ as the magnetic period of the undulator and $N$, the number of magnetic periods. The electron beam usually presents different horizontal and vertical beam emittances, but the undulator natural source size or divergence presents radial symmetry at the central cone (excluding polarization). The photon source size and divergence are approximately given by the convolution of the electron beam sizes and the undulator natural sizes. If we consider Gaussian distributions, the photon source is:
\begin{align}
\label{eq:photon big sigmas}
\Sigma_{h,v}=\sqrt{\sigma_{e(h,v)}^2 + \sigma_u^2} && \Sigma_{h,v}'=\sqrt{\sigma_{e(h,v)}'^2 + \sigma_u'^2}.
\end{align}

@@

Using the values for U18.3, $n=1$, and  $E=17225$~eV (Table \ref{tab:Undulators}) and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5~nm$^2$ (H$\times$V) for ESRF-1 and $10\times5$~nm$^2$ for EBS (Table~\ref{tab:HandCalculations}). The horizontal focal spot for the ESRF-1 lattice is larger than the specified target value for the beamline, therefore the VSS could be closed to 50~$\mu$m thus reducing the beam size at the VSS plane by a factor ca. 8.
%\inred{8 $\approx$ 403.8/50=8.076}
Therefore, for ESRF-1, the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5~nm$^2$.

@@

\begin{table}\label{tab:HandCalculations}
    \centering
    \caption{Photon beam size and divergence for selected ID16A beamline positions. Values below were obtained considering undulator U18.3 tuned to its first harmonic at the photon energy of $E=17225$~eV for both the high-$\beta$ (ESRF-1) and the EBS straight sections. Values are FWHM. Values in parentheses correspond to closing the VSS slit to 50~$\mu$m (21.3 if approximating by the Gaussian $\sigma$).}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{r|ccc|ccc}
                            &               & \textbf{ESRF (high-$\beta$)} &   &         & \textbf{EBS}  &    \\ 
    Element:                & Undulator     & VSS          & Sample            & Undulator & VVS   & Sample       \\
    Position [m]:           & 0.00          & 40.00        & 185.00            & 0.00    & 40.00   & 185.00       \\
    FWHM$_h$ [$\mu$m]:      & 977.19        & 404.00 (50.0)& 0.139 (0.017)     & 71.26   & 29.46   & 0.0102       \\
    FWHM$_v$ [$\mu$m]:      & 9.60          & 482          & 0.0052            & 10.02   & 486.5   & 0.0054       \\
    \newline
    FWHM'$_h$ [$\mu$rad]:   & 26.96         & 65.22        & 189062   & 15.60   & 37.74   & 109404       \\
    FWHM'$_v$ [$\mu$rad]:   & 12.04         & 12.04                 & 22264    & 12.16   & 12.16   & 22489        \\                                                   
    \end{tabular}%
    }
\end{table}

Next, we want to estimate the number of photons in the sample. For that, we need the spectrum at the source. It is not straightforward to get this information analytically but one can calculate the flux and power using different codes. Perhaps the most advanced and well-maintained codes used in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}. The XOPPY package in the OASYS suite implements three codes: US \cite{codeUS}, URGENT \cite{codeURGENT} and SRW. Figure~\ref{fig:FluxU18} shows the spectra calculated using SRW over an aperture of $1 \times 1$~mm$^2$ placed at $d=27.1$~m. This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use ($\max(\Sigma') \times d = 0.73$~mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}. The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and $1.9\cdot10^{14}$ for ESRF-1 (see \ref{fig:FluxU18}). One can appreciate how the emittance plays an important role in both peak width and especially in peak value. The number of photons per second after the multilayer monochromator is $2.9\cdot10^{15}$ for EBS and $1.9\cdot10^{15}$ for ESRF-1, considering that the energy bandwidth transmitted by a multilayer monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in Fig.~\ref{fig:FluxU18} is 10$^{-3}$, and supposing, at this point, ideal reflectivity (100~\%) for the multilayers.

\begin{figure}\label{fig:FluxU18}
    \centering
    \includegraphics[width=0.8\textwidth]{GRAPHICS/fluxU18.png}
    \caption{Flux of the U18.3 undulator integrated over a square slit placed at 27.1~m from the ID centre with aperture of 1 and 4 mm.}
\end{figure}

To estimate the geometrical transmission of the beamline we have to consider which optical elements crop the beam and remove photons. This can be analyzed using their angular acceptance or numerical aperture. The multilayer monochromator accepts the whole beam. We consider a VSS closed to $S_h=50~\mu$m in horizontal and fully opened in vertical. For ESRF-1 its transmission is 1/8 (as discussed before) and 1 for EBS. In a similar way, the KB mirrors will crop the beam to a size that is the projection of their useful optical length onto a direction perpendicular to the optical axis. Using a mirror length of $L_v=60$~mm and $L_h=26$~mm for VFM and HFM, respectively, and incident angle of 15~mrad, the KB behaves as an aperture of $900 \times 390~\mu$m$^2$. We get N.A. values of $6.2\times21.6~\mu$rad$^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V). The transmission coefficient is the ratio of the accepted N.A. over the beam divergence, resulting in 9.5\% (H) and 0.18\% (V)for ESRF-1 and 16.3\% and 0.18\% for EBS. The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF-1 and 2.95\% for EBS, therefore $3.8\cdot10^{12}$ photons/s for ESRF-1 and $8.5\cdot10^{13}$ photons/s for EBS. 
%\todo{check with references $10^9$ at 33.6 keV. I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12. How closed are the slits? We used the undulator peak value over the whole ML bandwidth, that overestimates the integrated flux.}
%\todo{need to consider ML reflectivity for the 3 reflections, also undulator is not perfect, the horizontal slit may be closed to less than $50 \mu m$, also KB mirrors may not be fully illuminated and slope error losses due to initial ML mono are neglected}

@@

To assess the focused beam size given by a coherent beam, one can consider the beamline in two sections: First, the ML monochromator and VSS that produce a coherent plane wave as the source (vertical) and VSS (horizontal) are both very far from the entrance aperture of the focusing optics. Second, the KB-system playing a dual role, acting as a focusing element of focal distance $f=5$~cm, but also cropping the beam by acting as a square aperture of $D \times D=$ 400$\times$400~$\mu$m$^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation).
%\todo{REMOVE: we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstream of the sample, and 390 in V approximated by 400}.
This aperture has two effects: i) as discussed before, it removes a lot of off-axis photons increasing the CF to a value close to one, and ii) broadens of the focal spot because of coherent diffraction. The diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$,
%\todo{include citation?}
where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate. Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75~nm$^2$ for the focal spot. This spot size is of the order or larger than that calculated geometrical considerations alone and therefore contributes significantly to the overall size of the focused beam. One can conclude that for the conditions in use this is a diffraction-limited beamline.

@@

%-------------------------------------------------------------------------
% 
%-------------------------------------------------------------------------
\subsection{Ray-tracing calculations}
\label{level1}
In this section, we perform ray-tracing calculations to study the effect of real optics (including aberrations and slope errors) upon the beam size and beamline transmission. The first subsection deals with the effect of aberrations and transmission due to geometrical considerations. The second subsection shows how the effect of slope errors of the KB mirrors degrade the focal spot.

Ray-tracing is a simulation method based on tracing some light rays along the optical system from the source to the image and retrieving the statistics of rays at the image to measure beam characteristics. For simple systems one can perform ray-tracing even by hand, selecting some principal rays that will define the location and envelope of the image. Using computers one can trace thousands of rays and calculate the position and divergence distribution of the rays using statistics. Rays are usually generated by Monte Carlo sampling the source characteristics. Ray-tracing is based on geometrical optics: a ray is a solution of the Helmholtz equation and travels in vacuum along a straight line. Rays are intercepted by the optical elements, which change their trajectory. For reflectors, this change of direction is given by the specular reflection laws. For refractors, the change of direction is defined by  Snell's law. Because Snell's law uses the refractive index which is wavelength-dependent, it is then interesting to assign additional attributes to each ray, such as the wavelength. This will permit us to calculate the refractive index that should be used for each ray. The characteristics of each ray can also be extended to include the electric field amplitude. This allows extending the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmission. This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combinations of them (e.g., KBs); lenses, compound refractive lenses, and transfocators; gratings of any type and shape, crystal systems, etc. Ray-tracing is a simple and extremely powerful technique for calculating the main characteristics of the photon beam (size, divergence, photon distribution) at every point of the beamline. Although very powerful commercial ray-tracing packages are available in the market, they are generally ill-adapted for synchrotron simulations because they do not cover the particularities of X-ray devices and sources. For that purposes, several packages created in the synchrotron community and are available, like SHADOW \cite{codeSHADOW}, RAY \cite{codeRAY}, McXtrace \cite{codeMCXTRACE} or XRT \cite{codeXRT}. For this study, we use SHADOW and its interface ShadowOUI \cite{codeSHADOWOUI}, available as a module of the OASYS suite.

We performed ray-tracing of the ID16A beamline using an undulator source tuned to have its resonance at 17225~eV, considering the ML as a cylindrical mirror, and using elliptical cylinder reflectors with finite dimensions as the KB focusing elements. The reflector shape is chosen to be elliptical as this is the theoretical shape producing a point-to-point focusing. For a first calculation, the reflectors (ML and KBs) are considered perfect and have no slope errors. 
The results in Fig.~\ref{fig:ray-tracing}a show two phenomena. One is that the migration from ESRF-1 lattice to the EBS will significantly improve the horizontal size, and the VSS is not needed. For EBS, the whole beam passes though the VSS, producing a final spot with almost Gaussian horizontal intensity profile instead of the step distribution for ESRF-1. Secondly, one can observe a much larger luminosity in the EBS case. The transmission values are 0.162\% for ESRF-1 and 2.22\% for EBS. These values confirm the analytical estimations (0.2\% for ESRF-1 and 2.95\% for EBS). As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF-1 and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore it is expected some broadening of the beam due to diffraction effects originated by the mirrors acceptance.

\begin{figure}
\label{fig:ray-tracing}
\centering
~~~~~ESRF-1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~EBS
\newline
\includegraphics[width=0.45\textwidth]{GRAPHICS/idealelementsESRF.png}
\includegraphics[width=0.45\textwidth]{GRAPHICS/idealelementsEBS.png}

\includegraphics[width=0.45\textwidth]{GRAPHICS/hybridnoerrorsESRF.png}
\includegraphics[width=0.45\textwidth]{GRAPHICS/hybridnoerrorsEBS.png}

\includegraphics[width=0.45\textwidth]{GRAPHICS/hybriderrorsESRF.png}
\includegraphics[width=0.45\textwidth]{GRAPHICS/hybriderrorsEBS.png}
\caption{Image produced by different types of ideal focusing for the ESRF (left column) and EBS (right column). The source with parameters in Table~\ref{tab:HandCalculations}. 
The first row has been calculated using the ideal reflectors (cylindrical for ML, elliptical for the BK mirrors) without slope errors. 
The second row shows the image calculated the hybrid option (ray-tracing plus coherent diffraction) also for ideal reflectors with no slope errors. It shows the focal spot degradation due to the diffraction arising from the beam aperturing by the KB mirrors. The calculations in the bottom row also include slope errors using the real mirror profiles measures at the ESRF metrology lab (see Appendix~\ref{appendix:metrology}).  Further beam broadening and the growth of satellite structures in the vertical intensity distribution can be observed. 
The focal beam dimensions (FWHM of the histograms) are written in the plot title.
}
\end{figure}

@@

The surface figure of the optical elements is another limiting effect for determining the size and intensity of the beam at the image position. It is possible to infer the effect of mirror imperfections and in particular slope errors by only knowing some statistical parameters such as the RMS slope error and then building synthetic surface profiles to be used in the simulations. It is, though, highly recommended to use metrology data from real mirrors, when possible. When not available, one can use data from existing mirrors from a database (e.g., \cite{dabam}) and extrapolate some of the effects to the hypothetical new mirrors. This is the typical approach used when designing a new beamline where the mirrors do not exist yet. However, in our case, the mirrors already exist and have been extensively characterized at the ESRF optical metrology laboratory.  The metrology data of the mirrors for the beamline under study are presented in Appendix~\ref{appendix:metrology}. These experimental data are used in all the simulations in this paper. Simulations using the hybrid model including slope errors are shown in Fig.~\ref{fig:ray-tracing}c. A small broadening of the peak due to slope errors can be observed, but, more importantly, a new satellite structure close to the vertical intensity peak is produced.

>>>>

\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{color,xcolor}
%\usepackage{siunitx}
\usepackage[normalem]{ulem}

@@

\begin{abstract}
We consider different approaches to simulate a modern X-ray beamline. Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performance. Parameters such as flux, dimensions and intensity distribution of the focused beam and coherence properties are obtained from simple analytical calculations to sophisticated computer simulations using ray-tracing and wave optics techniques. A latest-generation X-ray nanofocusing beamline for coherent applications (ID16A at the ESRF) has been chosen to study in detail the issues related to highly demagnifying synchrotron sources and exploiting the beam coherence. The performance of the beamline is studied for two storage rings: the old ESRF-1 (emittance 4000~pm) and the new ESRF-EBS (emittance 150~pm). In addition to traditional results in terms of flux and beam sizes, an innovative study on the partial coherence properties based on the propagation of coherent modes is presented. The different algorithms and methodologies are implemented in the software suite OASYS \cite{codeOASYS}. Those are discussed with emphasis placed upon the their benefits and limitations of each.
\end{abstract}

@@

Accurate calculation and quantitative evaluation of the parameters related to X-ray optics are of great importance for designing, building and exploiting new beamlines. Every modern beamline follows a procedure of conceptual design plus technical design where detailed simulations are essential. These calculations allow testing of the design parameters in a virtual computer environment where advantages and limitations can be studied quantitatively. The selection of the optimal optical layout requires a detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optics. Deformations of the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformations usually done using finite element analysis. Traditionally the optics calculations for synchrotron beamlines are done using ray-tracing techniques. They model the X-ray beam as a collection of incoherent rays. This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc. However, the high coherence of the new sources can introduce the necessity to complement the ray-tracing methods with models including effects of diffraction and scattering from a coherent beam. A fully coherent hard X-ray beam is still a dream for the new generation storage rings. Despite a considerable increase in the coherent fraction of the new beams, the coherent fraction at typically hard X-ray photon energies is still a few percent for the new sources. Consequently, the ray-tracing analysis does not become obsolete but has to be complemented with an analysis based on wave optics. Because of the partial coherence, the wave optics analysis becomes nontrivial and can be done with different levels of approximation, as discussed in this work. In addition, wave optics simulations are usually computationally expensive.

To illustrate the applicability of a variety of methods and solutions for beamline optical simulations we have chosen the ESRF beamline ID16A~\cite{ID16A}. This beamline provides a high-brilliance beam focused to a few nanometers. It displays characteristics increasingly aimed for in modern beamlines: extremely high demagnification (sub-micrometer focused beam size) and high coherence. It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state~\cite{XRFID16A}.

@@

 In this paper, some quantitative results on how the storage ring upgrade will affect a particular beamline are presented. The paper is structured as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and explain the particular choice of the parameters used in the calculations. Section~\ref{Simulations} is the central part of the paper, and describes the different methods used to analyze the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline. In Section \ref{level0} we calculate with simple methods approximate values of focal size, flux and coherent fraction. Accuracy can be improved using geometrical methods in Section \ref{level1} that allow a more complete simulation of the beamline optics within certain limits which are related, in this example, to high demagnification and high coherence of the studied beamline. The wave optics section~\ref{level2} includes different models of increasing complexity as the required accuracy of the results becomes more demanding. In the Discussion (Section \ref{discussion}) the results of the different simulation methods are commented. The hierarchical approach allows more accurate and better results but the price to pay is to deal with more complex and computationally expensive tools. Finally, Section \ref{summary} summarizes the results.

@@

The undulator natural photon source size and divergence are given by \cite{elleaume}:
\begin{align}
    \label{eq:photon small sigmas}
    \sigma_u=\frac{2.74}{4\pi}\sqrt{\lambda L}\approx \frac{1}{2  \pi}\sqrt{2 \lambda L}  && \sigma_u' = 0.69\sqrt{\frac{\lambda}{L}}\approx \sqrt{\frac{\lambda}{2 L}},
\end{align}
where $\lambda$ is the emitted wavelength and $L=N\Lambda_u$ is the undulator length, with $\Lambda_u$ as the magnetic period of the undulator and $N$, the number of magnetic periods. The electron beam usually presents different horizontal and vertical beam emittances, but the undulator natural source size or divergence presents radial symmetry at the central cone (excluding polarization). The photon source size and divergence are approximately given by the convolution of the electron beam sizes and the undulator natural sizes. If we consider Gaussian distributions, the photon source has sizes $\Sigma$ and divergences $\Sigma'$ given by:
\begin{align}
\label{eq:photon big sigmas}
\Sigma_{h,v}=\sqrt{\sigma_{e(h,v)}^2 + \sigma_u^2} && \Sigma_{h,v}'=\sqrt{\sigma_{e(h,v)}'^2 + \sigma_u'^2}.
\end{align}

@@

Using the values for U18.3, $n$=1, and  $E$=17225~eV (Table \ref{tab:Undulators}) and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5~nm$^2$ (H$\times$V) for ESRF-1 and $10\times5$~nm$^2$ for EBS (Table~\ref{tab:HandCalculations}). The horizontal focal spot for the ESRF-1 lattice is larger than the specified target value for the beamline, therefore the VSS could be closed to 50~$\mu$m thus reducing the beam size at the VSS plane by a factor ca. 8.
%\inred{8 $\approx$ 403.8/50=8.076}
Therefore, for ESRF-1, the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5~nm$^2$.

@@

\begin{table}\label{tab:HandCalculations}
    \centering
    \caption{Photon beam size and divergence for selected ID16A beamline positions. Values below were obtained considering undulator U18.3 tuned to its first harmonic at the photon energy of $E=17225$~eV for both the high-$\beta$ (ESRF-1) and the EBS straight sections. Values are FWHM. Values in parentheses correspond to closing the VSS slit to 50~$\mu$m.
    %(21.3 if approximating by the Gaussian $\sigma$).
    }
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{r|ccc|ccc}
                            &               & \textbf{ESRF-1 (high-$\beta$)} &   &         & \textbf{EBS}  &    \\ 
    Element:                & Undulator     & VSS          & Sample            & Undulator & VVS   & Sample       \\
    Position [m]:           & 0.00          & 40.00        & 185.00            & 0.00    & 40.00   & 185.00       \\
    FWHM$_h$ [$\mu$m]:      & 977.19        & 404.00 (50.0)& 0.139 (0.017)     & 71.26   & 29.46   & 0.0102       \\
    FWHM$_v$ [$\mu$m]:      & 9.60          & 482          & 0.0052            & 10.02   & 486.5   & 0.0054       \\
    \newline
    FWHM'$_h$ [$\mu$rad]:   & 26.96         & 65.22        & 189062   & 15.60   & 37.74   & 109404       \\
    FWHM'$_v$ [$\mu$rad]:   & 12.04         & 12.04                 & 22264    & 12.16   & 12.16   & 22489        \\                                                   
    \end{tabular}%
    }
\end{table}

Next, we want to estimate the number of photons in the sample. For that, we need the spectrum at the source. It is not straightforward to get this information analytically but one can calculate the flux and power using different codes. Perhaps the most advanced and well-maintained codes used in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}. The XOPPY package in the OASYS suite implements three codes: US \cite{codeUS}, URGENT \cite{codeURGENT} and SRW. Figure~\ref{fig:FluxU18} shows the spectra calculated using SRW over an aperture of $1 \times 1$~mm$^2$ placed at $d=27.1$~m. This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use.
%($\max(\Sigma') \times d = 0.73$~mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}. 
The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and $1.9\cdot10^{14}$ for ESRF-1 (see \ref{fig:FluxU18}). One can appreciate how the emittance plays an important role in both peak width and especially in peak value. The number of photons per second after the multilayer monochromator is $2.9\cdot10^{15}$ for EBS and $1.9\cdot10^{15}$ for ESRF-1, considering that the energy bandwidth transmitted by a multilayer monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in Fig.~\ref{fig:FluxU18} is 10$^{-3}$, and supposing, at this point, ideal reflectivity (100~\%) for the multilayers.

\begin{figure}\label{fig:FluxU18}
    \centering
    \includegraphics[width=0.95\textwidth]{GRAPHICS/fluxU18.png}
    \caption{Flux of the U18.3 undulator integrated over a square slit placed at 27.1~m from the ID centre with aperture of 1 and 4 mm.}
\end{figure}

To estimate the geometrical transmission of the beamline we have to consider which optical elements crop the beam and remove photons. This can be analyzed using their angular acceptance or numerical aperture. The multilayer monochromator accepts the whole beam. We consider a VSS closed to $S_h$=50~$\mu$m in horizontal and fully opened in vertical. For ESRF-1 its transmission is 1/8 (as discussed before) and 1 for EBS. In a similar way, the KB mirrors will crop the beam to a size that is the projection of their useful optical length onto a direction perpendicular to the optical axis. Using a mirror length of $L_v$=60~mm and $L_h$=26~mm for VFM and HFM, respectively, and incident angle of 15~mrad, the KB behaves as an aperture of 900 $\times$ 390~$\mu$m$^2$. We get N.A. values of 6.2 $\times$ 21.6~$\mu$rad$^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V). The transmission coefficient is the ratio of the accepted N.A. over the beam divergence, resulting in 9.5\% (H) and 0.18\% (V)for ESRF-1 and 16.3\% and 0.18\% for EBS. The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF-1 and 2.95\% for EBS, therefore $3.8\cdot10^{12}$ photons/s for ESRF-1 and $8.5\cdot10^{13}$ photons/s for EBS. 
%\todo{check with references $10^9$ at 33.6 keV. I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12. How closed are the slits? We used the undulator peak value over the whole ML bandwidth, that overestimates the integrated flux.}
%\todo{need to consider ML reflectivity for the 3 reflections, also undulator is not perfect, the horizontal slit may be closed to less than $50 \mu m$, also KB mirrors may not be fully illuminated and slope error losses due to initial ML mono are neglected}

@@

To assess the focused beam size given by a coherent beam, one can consider the beamline in two sections: First, the ML monochromator and VSS that produce a coherent plane wave as the source (vertical) and VSS (horizontal) are both very far from the entrance aperture of the focusing optics. Second, the KB-system playing a dual role, acting as a focusing element of focal distance $f=5$~cm, but also cropping the beam by acting as a square aperture of roughly D$\times$D=400$\times$400~$\mu$m$^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation).
%\todo{REMOVE: we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstream of the sample, and 390 in V approximated by 400}.
This aperture has two effects: i) as discussed before, it removes a lot of off-axis photons increasing the CF to a value close to one, and ii) broadens of the focal spot because of coherent diffraction. The diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$,
%\todo{include citation?}
where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate. Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75~nm$^2$ for the focal spot. This spot size is of the order or larger than that calculated geometrical considerations alone and therefore contributes significantly to the overall size of the focused beam. One can conclude that for the conditions in use this is a diffraction-limited beamline.

@@

%-------------------------------------------------------------------------
% 
%-------------------------------------------------------------------------
\subsection{Ray-tracing calculations}
\label{level1}
In this section, we perform ray-tracing calculations to study the effect of real optics (including aberrations and slope errors) upon the beam size and beamline transmission. The first part deals with the effect of aberrations and transmission due to geometrical considerations. The second part shows how the effect of slope errors of the KB mirrors degrade the focal spot.

Ray-tracing is a simulation method based on tracing some light rays along the optical system from the source to the image and retrieving the statistics of rays at the image to measure beam characteristics. For simple systems one can perform ray-tracing even by hand, selecting some principal rays that will define the location and envelope of the image. Using computers one can trace thousands of rays and calculate the position and divergence distribution of the rays using statistics. Rays are usually generated by Monte Carlo sampling the source characteristics. Ray-tracing is based on geometrical optics: a ray is a solution of the Helmholtz equation and travels in vacuum along a straight line. Rays are intercepted by the optical elements, which change their trajectory. For reflectors, this change of direction is given by the specular reflection laws. For refractors, the change of direction is defined by  Snell's law. Because Snell's law uses the refractive index which is wavelength-dependent, it is then interesting to assign additional attributes to each ray, such as the wavelength. This will permit us to calculate the refractive index that should be used for each ray. The characteristics of each ray can also be extended to include the electric field amplitude. This allows extending the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmission. This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combinations of them (e.g., KBs); lenses, compound refractive lenses, and transfocators; gratings of any type and shape, crystal systems, etc. Ray-tracing is a simple and extremely powerful technique for calculating the main characteristics of the photon beam (size, divergence, photon distribution) at every point of the beamline. 
%Although very powerful commercial ray-tracing packages are available in the market, they are generally ill-adapted for synchrotron simulations because they do not cover the particularities of X-ray devices and sources. For that purposes, 
Several packages created in the synchrotron community and are available, like SHADOW \cite{codeSHADOW}, RAY \cite{codeRAY}, McXtrace \cite{codeMCXTRACE} or XRT \cite{codeXRT}. For this study, we use SHADOW and its interface ShadowOUI \cite{codeSHADOWOUI}, available as a module of the OASYS suite.

We performed ray-tracing of the ID16A beamline using an undulator source tuned to have its resonance at 17225~eV and approximated using Gaussian distributions (Eqs.~\ref{eq:photon big sigmas}), considering the ML as a cylindrical mirror, and using elliptical cylinder reflectors with finite dimensions as the KB focusing elements. The KB reflector shape are elliptical as required for a point-to-point focusing. For a first calculation, the reflectors (ML and KBs) are considered perfect and have no slope errors. 
The results in Fig.~\ref{fig:ray-tracing}a show two phenomena. One is that the migration from ESRF-1 lattice to the EBS will significantly improve the horizontal size, and the VSS is not needed. For EBS, the whole beam passes though the VSS, producing a final spot with almost Gaussian horizontal intensity profile instead of the step distribution for ESRF-1. Secondly, one can observe a much larger luminosity in the EBS case. The transmission values are 0.16\% for ESRF-1 and 2.22\% for EBS. These values confirm the analytical estimations (0.2\% for ESRF-1 and 2.95\% for EBS). As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF-1 and 2.73\% for EBS), meaning that a very high coherence of the beam is expected. It is therefore expected some broadening of the beam due to diffraction effects originated by the mirrors acceptance.

@@

\begin{figure}
\label{fig:ray-tracing}
\centering
~~~~~ESRF-1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~EBS
\newline
\includegraphics[width=0.95\textwidth]{GRAPHICS/figure3.png}
\caption{Image produced by different types of ideal focusing for the ESRF (left column) and EBS (right column). The source follows Gaussian distributions with parameters in Table~\ref{tab:HandCalculations}. 
a) Simulations using the ideal reflectors (cylindrical for ML, elliptical for the BK mirrors) without slope errors. 
b) Images calculated using the hybrid method (ray-tracing plus coherent diffraction) with ideal reflectors with no slope errors. It shows the focal spot degradation due to the diffraction arising from the cropping of the beam by KB mirrors. c) Simulations using the hybrid method including slope errors from the real mirror profiles measured at the ESRF metrology lab (see Appendix~\ref{appendix:metrology}). It can be observed further beam broadening and the growth of satellite structures. 
The focal beam dimensions (FWHM of the histograms) are written in the plot title.
}
\end{figure}
